---
title: "p8105_hw2_zc2555"
author: "Ziyu Chen"
date: "10/7/2021"
output: github_document
---

## R set up
```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Question 1 
_Mr.Trash Wheel Sheet clean up_

```{r message=FALSE}
trash_df = read_excel("./Trash_data.xlsx", 1) %>%
   janitor::clean_names() %>%
   select(-c(x15, x16, x17)) %>%
   drop_na() %>%
   mutate(sports_balls = round(sports_balls, digits = 0)) 
```


_Precipitation 2019 data cleaning_

```{r message=FALSE}
precipitation_df_2019 = read_excel("./Trash_data.xlsx", 6, skip = 1) %>%
  drop_na()  %>%
add_column(Year = 2019)
```

_Precipitation 2018 data cleaning_

```{r message = FALSE}
precipitation_df_2018 = read_excel("./Trash_data.xlsx", 7, skip = 1) %>%
  drop_na() %>%
  add_column(Year = 2018)
```


_merge 2019 and 2018 together_
```{r message = FALSE}
precipitation_merge = rbind(precipitation_df_2018, precipitation_df_2019)
```

_The following paragraph describe the details about 3 datasets_

* 
The trash_df dataset contains `r nrow(trash_df)` rows and `r ncol(trash_df)` columns.
The median number of sports ball is `r median(pull(trash_df, sports_balls))`

*
The precipitation_df_2018 dataset contains `r nrow(precipitation_df_2018)` rows and `r ncol(precipitation_df_2018)` columns. The total precipitation in 2018 is `r sum(pull(precipitation_df_2018,Total)) `

*
The precipitation_df_2019 dataset contains `r nrow(precipitation_df_2019)` rows and `r ncol(precipitation_df_2019)` columns. The recorded total precipitation in 2018 is only `r sum(pull(precipitation_df_2019,Total)) `. 


## Question2

_import cvs files_
_clean pols_month.csv_

```{r}
pols_month = read_csv(file = "./fivethirtyeight_datasets/pols-month.csv")
pols_month_separated = separate(
  pols_month,
  mon,
  c("Year", "Month", "Date" )) %>%
  select(-"Date")
pols_month_separated =
  pols_month_separated %>% mutate(Month = month.name[as.integer(Month)],
                                  Year = as.integer(Year)) 
  
pols_month_separated = pols_month_separated %>%
  mutate(president = case_when(
    prez_dem == 1 ~ 'dem',
    prez_dem == 0 ~ 'gop'
  )) %>%
  select(-c("prez_dem","prez_gop"))


```
_clean snp.csv_

```{r}
snp_df = read_csv(file = "./fivethirtyeight_datasets/snp.csv")
snp_df_separated = separate(
  snp_df,
  date,
  c("Month", "Date", "Year")) %>%
  select(-"Date")
snp_df_separated =
  snp_df_separated %>% mutate(Month = month.name[as.integer(Month)],
                              Year = as.integer(Year),
                              Year = ifelse(Year<= 21, Year+2000, Year+1900)
                              )
  
snp_df_separated = snp_df_separated %>% select (Year, Month, close)
```

_import unemployment.csv_


```{r}
unemployment_df = read_csv(file = "./fivethirtyeight_datasets/unemployment.csv")

unemployment_tidy_df= 
  pivot_longer(
    unemployment_df,
    Jan:Dec,
    names_to = "Month",
    values_to = "rate"
  ) %>% mutate(Year = as.integer(Year))
```


_merging 3 datasets into one_

```{r}
merged_df=left_join(pols_month_separated, snp_df_separated, by=c("Year", "Month"))

merged_all=left_join(merged_df, unemployment_tidy_df,by=c("Year", "Month"))
```

*Paragraph about three datasets

## Question 3
_cleaning up the data_

```{r}
name_df = read_csv(file = "./Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
  childs_first_name = str_to_sentence(childs_first_name),
  ethnicity = str_to_sentence(ethnicity),
  ethnicity = recode ( ethnicity,
                       'Asian and paci' = "Asian and pacific islander",
                       'Black non hisp' = "Black non hispanic",
                       'White non hisp' = "White non hispanic") )%>%
  distinct()
  



```

```{r}
subset(name_df, childs_first_name == 'Olivia')%>%pivot_wider(
            names_from = 'ethnicity',
            values_from = 'rank'
)
```



